---
title: "FInalproject"
author: "Emilia Korobowicz"
date: "31/07/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(cache = TRUE)
```

```{r}
d_fun <- function(id, df) {
    tmp <- subset(df, pidp == id)
    res <- tmp[seq(2, nrow(tmp), 1), "vote6"] - tmp[seq(1, nrow(tmp)-1, 1), "vote6"]
    names(res) <- c("0_1", "1_2", "2_3")
    return(res)
}
```



Load libraries

```{r}
library(tidyverse)
library(data.table)
library(ggplot2)
library(data.table)
library(tibble)
```

I will explore political interest (*vote6*) and how it changes over the political cycle.I will try to combine population level analysis (i.e. mean political interest) with individual level analysis.
General elections: 2010, 2015 and 2017
Wave 1 - 2009-2011 -> Around election-election (0)
Wave 2 - 2010-2012 -> After election (+1)
Wave 3 - 2011-2013 -> After election (+2)
Wave 4- 2012-2014 -> After election (+3)
Wave 5 - 2013-2015 -> Around election (0)
Wave 6 - 2015-2017 -> Around election (0)
Wave 7 - 2017-2019 -> After election (+1)


First I will read the data. I will want to read and join the data for the first 7 waves of the Understanding Society. Wave 8 does not have a variable for political interest). I will keep 5 variables: personal identifier, sample origin, sex, age and political interest.

```{r}
files <- dir("/Users/emiliakorobowicz/Desktop/DataScience3/EmilysRepo/NewRepo/Data/UKDA-6614-tab/tab/",
             pattern = "indresp",
             recursive = TRUE,
             full.names = TRUE)

files <- files[stringr::str_detect(files, "ukhls")]
files

vars <- c("memorig", "sex_dv", "age_dv", "vote6")

for (i in 1:7) {
        
        varsToSelect <- paste(letters[i], vars, sep = "_")
       
        varsToSelect <- c("pidp", varsToSelect)
      
        data <- fread(files[i], select = varsToSelect)
        if (i == 1) {
                all7 <- data  
        }
        else {
                all7 <- full_join(all7, data, by = "pidp")
        }
        
        rm(data)
} 

```



```{r}
Long <- all7 %>% reshape(idvar = "pidp", direction = "long", 
                         v.names = c("memorig", "sex_dv", "age_dv", "vote6"),
                         times = letters[1:7],
                         varying = list(grep("memorig", names(all7), value = TRUE), 
                                        grep("sex_dv", names(all7), value = TRUE),
                                        grep("age_dv", names(all7), value = TRUE), 
                                        grep("vote6", names(all7), value = TRUE)))

Long
```

Filter and recode.
Now I will filter the data keeping only respondents from the original UKHLS sample for Great Britain (memorig == 1). I also want to clean the variables for sex (recoding it to "male" or "female") and political interest (keeping the values from 1).
## to 4 and coding all negative values as missing)

```{r}
Long <- Long %>% 
    filter(memorig == 1) %>%
    mutate(sex_dv = recode(sex_dv, "1" = "Male", "2" = "Female")) %>% 
    mutate(vote6 = ifelse(vote6 < 0, NA, vote6)) %>% 
    mutate(time_cycle = ifelse(time %in% c("a", "e", "f"), 0, 
                               ifelse(time %in% c("b", "g"), 1, 
                                      ifelse(time == "c", 2, 3
                                             )
                                      )
                               )
           )

```


Calculate mean political interest by sex, time and time_cycle

```{r}
meanVote6 <- subset(Long, !is.na(sex_dv)) %>% group_by(sex_dv, time)  %>% summarise(mean = mean(na.omit(vote6)))
meanVote6
meanVote6_b <- subset(Long, !is.na(sex_dv)) %>% group_by(sex_dv, time_cycle)  %>% summarise(mean = mean(na.omit(vote6)))
meanVote6_b
```

First plot - MEAN political interest vs time_cycle by gender. Notice that here we are looking at changes in MEAN political interest, rather than individual trends.

In this plot, we can see that the trends in female and in males are very similar; although the mean political interest is persistently higher in females than in males (~0.33 units).

The mean political interest progresively declines after the elections, reaching a minimum about 1 year after the election. Then, there is a rapid increase in political interest.
Until second year after the elections; afterwhich there still an increasing tren but it is a lot more stable (there is a kind of plateau).

```{r}
ggplot(meanVote6_b, aes(time_cycle, mean, group = sex_dv, color = sex_dv)) + geom_point() +
    geom_smooth(size = 0.5) + theme_bw() + ylab("Mean political interest") + xlab("Time (years after election)") +
    scale_color_manual(values = c("coral", "darkolivegreen4")) +  labs(color = "Sex") + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

I calculated the mean difference between males and females [within waves].
```{r}
meanVote6_b_reshaped <- as.data.frame(meanVote6_b) %>% reshape(idvar = "sex_dv", timevar = "time_cycle", direction = "wide")
meanVote6_b_reshaped[1, -1] - meanVote6_b_reshaped[2, -1] 
```

Estimate stability of political interest over the political cycle.

To simplify interpretation, I will consider only wave a-d (timepoints 0, 1, 2, 3), and will only keep only the respondents with non-missing values for political interest in these 4 waves
I will do this analysis at the individual level, rather than taking the mean(vote6) of each wave as we did beforehand.

```{r}
tmp <- subset(Long, time %in% c("a", "b", "c", "d"))
respondents_NA <- tmp %>% group_by(pidp) %>% summarise(no_missing = sum(!is.na(vote6)))
keep_respondents <- subset(respondents_NA, no_missing == 4) ## complete observations
df <- subset(tmp, pidp %in% keep_respondents$pidp &
                   !is.na(sex_dv) & !is.na(age_dv))
n_participants_dropped <- length(unique(respondents_NA)$pidp)
n_participants_dropped # 48111 # these numbers should be mentioned in the report
n_participants_included <- length(unique(keep_respondents)$pidp) 
n_participants_included # 21060 # these numbers should be mentioned in the report
df <- arrange(df, time) # make sure df is sorted by time
```

## Calculate Delta by time_cycle for each person

```{r}
female_delta <-do.call(rbind, lapply(unique(subset(df, sex_dv == "Female")$pidp), d_fun, df))
female_delta <- as.data.frame(female_delta)
rownames(female_delta) <- unique(subset(df, sex_dv == "Female")$pidp)
male_delta <-do.call(rbind, lapply(unique(subset(df, sex_dv == "Male")$pidp), d_fun, df))
male_delta <- as.data.frame(male_delta)
rownames(male_delta) <- unique(subset(df, sex_dv == "Male")$pidp)
head(female_delta)
head(male_delta)
```
here 0_1 is the change from wave a to wave b; 1_2 change from wave b to wave c; and 2_3 change from wave c to wave d
each row corresponds to a participants, so basically female_delta or male_delta, show the change
in political interest of each individual along the political cycle
I will tabulate the change in political interest for each of the 3 intervals (0_1), (1_2), (2_3)

1) Analysis for females

```{r}
fd_summary <- as.data.frame(do.call(rbind, lapply(female_delta, function(x) 100*(table(x)/nrow(female_delta)))))
fd_summary <- round(fd_summary, 2) 
fd_summary
```
Each row of fd_summary represents the interval between two consecutive waves
The rows indicate the absolute change in political interest (vote 6). Each cell indicates
the percentage of participants that had a specific change in vote 6 for a given 
time interval (0_1, 1_2, 1_3).

Now I check the % of participants for which there was an increase in political interest for each of the 3 time intervals.

```{r}
rowSums(fd_summary[, 1:3])
```

Now I check the % of participants for which there was a decrease in political interest
for each of the 3 time intervals
```{r}
rowSums(fd_summary[, 5:7])
```

Key messages for each interval:
0_1
Overall
```{r}
mean(female_delta[, 1])
```

- 60% of participants did not change their political interest
- 19.8% of participants moved towards a higher value of vote6
- 19.6% of participants moved towards a lower value of vote6
However, on average there was a decrease of political interest (mean(female_delta[, 1])).
An overall decline in potical interest indicates that although there were more 
participants that gain political interest compared to those that lost political interest;
the individual drop in political interest was larger than the individual gain in political interest

1_2
Overall
```{r}
mean(female_delta[, 2])
```
- 62% of participants did not change their political interest
- 21.7% of participants moved towards a higher value of vote6
- 16.01% of participants moved towards a lower value of vote6
On average there was an increase of political interest mean(female_delta[, 2])

2_3
Overall
```{r}
mean(female_delta[, 3])
```
64% of participants did not change their political interest
19.2% of participants moved towards a higher value of vote6
17.01% of participants moved towards a lower value of vote6
In average there was an increase of political interest mean(female_delta[, 3])

Overall change 0_3
```{r}
table(rowSums(female_delta)/3)/nrow(female_delta) # 58% of participants did not change their political interest
sum((table(rowSums(female_delta)/3)/nrow(female_delta))[5:7]) # 24.71% moved to a higher value of vote6
sum((table(rowSums(female_delta)/3)/nrow(female_delta))[1:3]) # 17.34 moved to a lowe level of vote6

df_plot <-data.frame(delta = (rowSums(female_delta)/3))


ggplot(data = df_plot, aes(delta)) + geom_histogram(fill = "gray", color = "black") + theme_bw() + xlab("Delta a-d")
```

This distribution is quite symmetric, but slightly shifted to the right. This means that overall there is an increase in political interest through the political cycle


Now I plot delta trend
```{r}
female_delta_2 <- data.frame(delta = colSums(female_delta)/nrow(female_delta), time = c(0.5,1.5,2.5))
female_delta_2$sex = "Female"

ggplot(female_delta_2, aes(x = time, y = delta)) + geom_point(color = "coral") + geom_smooth(color = "coral") + xlab("Time (years after election)") +
    ylab("Delta") + theme_bw() + ylab("Count")
```
When we compare the change in political interest between consecutive years following an election, we can see that the largest change in politica interest occurs between the first and the second year.








2) Analysis for males
```{r}
md_summary <- as.data.frame(do.call(rbind, lapply(male_delta, function(x) 100*(table(x)/nrow(male_delta)))))
md_summary <- round(md_summary, 2) 
md_summary
```

Each row of fd_summary represents the interval between two consecutive waves. The rows indicate the absolute change in political interest (vote 6). Each cell indicates the percentage of participants that had a specific change in vote 6 for a given time interval (0_1, 1_2, 1_3)

Let's now check the % of participants for which there was an increase in political interest
for each of the 3 time intervals
```{r}
rowSums(md_summary[, 1:3])
```

Let's now check the % of participants for which there was a decrease in political interest
for each of the 3 time intervals
```{r}
rowSums(md_summary[, 5:7])
```

Key messages for each interval:
0_1
overall
```{r}
mean(male_delta[, 1])
```
- 60% of participants did not change their political interest
- 19.4% of participants moved towards a higher value of vote6
- 20.1% of participants moved towards a lower value of vote6
However, on average there was a decrease of political interest (mean(male_delta[, 1]))

1_2
overall
```{r}
mean(male_delta[, 2])
```
- 61% of participants did not change their political interest
- 17.7% of participants moved towards a higher value of vote6
- 16.01% of participants moved towards a lower value of vote6
On average there was an increase of political interest mean(female_delta[, 2])

2_3
overall
```{r}
mean(male_delta[, 3])
```
- 64% of participants did not change their political interest
- 19.3% of participants moved towards a higher value of vote6
- 17.1% of participants moved towards a lower value of vote6
On average there was an increase of political interest mean(male_delta[, 3])

Overall change 0_3
```{r}
table(rowSums(male_delta)/3)/nrow(male_delta) # 57% of participants did not change their political interest
sum((table(rowSums(male_delta)/3)/nrow(male_delta))[5:7]) # 23.3% moved to a higher value of vote6
sum((table(rowSums(male_delta)/3)/nrow(male_delta))[1:3]) # 19.7 moved to a lowe level of vote6



dm_plot <-data.frame(delta = (rowSums(male_delta)/3))
ggplot(data = df_plot, aes(delta)) + geom_histogram(fill = "gray", color = "black") + theme_bw() + xlab("Delta a-d")
```
This distribution is quite symmetric, but slightly shifted to the right. This means
that overall there is an increase in political interest through the political cycle

At this stage I will compare the the delta distributions for the two genders 
```{r}
library(cowplot)
m <- ggplot(data = dm_plot, aes(delta)) + geom_histogram(fill = "darkolivegreen4", color = "black", alpha = 0.8) +
    theme_bw() + xlab("Delta a-d") + ggtitle("Male") + ylab("Count") +
    scale_y_continuous(limits = c(0, 8500), breaks = seq(0, 10000, 1000))
f <- ggplot(data = df_plot, aes(delta)) + geom_histogram(fill = "coral", color = "black", alpha = 0.8) +
    theme_bw() + xlab("Delta a-d") + ggtitle("Female") + ylab('Count') +
    scale_y_continuous(limits = c(0, 7500), breaks = seq(0, 10000, 1000))
mf <- plot_grid(f, m, ncol = 2)
mf

```

Plot delta trend
```{r}
male_delta_2 <- data.frame(delta = colSums(male_delta)/nrow(male_delta), time = c(0.5,1.5,2.5))
male_delta_2$sex = "Male"

ggplot(male_delta_2, aes(x = time, y = delta)) + geom_point(color = "darkolivegreen4") + geom_smooth(color = "darkolivegreen4") + xlab("Time (years after election)") +
    ylab("Delta") + theme_bw()
```
When we compare the change in political interest between consecutive years following an election,
we can see that the largest change in political interest occurs between the first and the second year.

Plot delta trend combined for males and females
```{r}
male_female_delta <- rbind(female_delta_2, male_delta_2)
ggplot(male_female_delta, aes(x = time, y = delta, color = sex)) + geom_point() + geom_smooth() + xlab("Time (years after election)") +
    ylab("Delta") + theme_bw() + scale_color_manual(values = c("coral", "darkolivegreen4")) +
    labs(color = "Sex") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
    # legend.key = element_rect(fill = "white", color = "white"))
```

Compare changes between males and females
```{r}
rowSums(md_summary[, 1:3]) - rowSums(fd_summary[, 1:3]) 
wilcox.test(as.vector(as.matrix(md_summary)[, 1:3]), as.vector(as.matrix(fd_summary)[, 1:3]), paired = TRUE, alternative = "greater")
```
the percentage of cases with an overall drop in political interest is significantly 
higher in males than in females for all 3 time intervals considered
```{r}
rowSums(md_summary[, 5:7]) - rowSums(fd_summary[, 5:7])
```
there are no significant inter-gender differences for the percentage of cases with an overall increase in political interest
```{r}
wilcox.test(as.vector(as.matrix(md_summary)[, 5:7]), as.vector(as.matrix(fd_summary)[, 5:7]), paired = TRUE, alternative = "less")
```
